---
title: "HyPyP fNIRS coherence getting started in R"
author: "Patrice Fortin"
date: "2025-01-06"
output:
  pdf_document: default
  html_notebook: default
---

This R notebook shows how to load a CSV file produced by HyPyP-fNIRS from a `Cohort` object, and load it in R to do statistical analysis.

The `.csv` file here has been generated from the Jupyter Notebook `tutorial/fnirs_getting_started.ipynb`.

The code here is to help getting started in loading the transformed data and get it in a suitable dataframe format for analysis. The plots below are very basics and serve only as inspiration.

Here is a list of fields:

|Field | Description |
| - | ---- |
| **dyad**             | Label of the dyad |
| **is_intra**         | If this is an intra-subject coherence |
| **is_shuffle**       | If this dyad is a permutation for statistical analysis or a real dyad |
| **subject1**         | Label of the subject 1 |
| **subject2**         | Label of the subject 2 |
| **roi1**             | Label of the Region of Interest of subject 1 channel |
| **roi2**             | Label of the Region of Interest of subject 2 channel |
| **channel1**         | Label of the subject 1 channel |
| **channel2**         | Label of the subject 2 channel |
| **task**             | Label of the task |
| **epoch**            | Identifier of the epoch for this coherence |
| **section**          | Identifier of the section, if an epoch had to be splitted because of invalid data |
| **bin**              | Identifier of the bin, if the WTC has been sliced for balancing the weight of computed coherence |
| **coherence**        | Value of the coherence for this bin |
| **coherence_masked** | Ratio of the excluded data in this bin |
| **bin_time_range**   | Label of the time range for this bin |
| **bin_period_range** | Label of the period range for this bin (for grouping) |
| **wavelet_library**  | Name of the wavelet library used |
| **wavelet_name**     | Label of the specific wavelet (contains wavelet parameters) |

### Load libraries

```{r}
library("ggplot2")
```

### Load libraries and CSV file produced by `fnirs_cohort_example.ipynb`.

```{r}
csv_path <- "../data/results/fnirs_cohort_example.csv"
data <- read.csv(csv_path, header = TRUE, stringsAsFactors = FALSE)
```

### Cleanup loaded data and have factors for better use with ggplot

```{r}
data <- within(data, {
  dyad <- as.factor(dyad)
  subject1 <- as.factor(subject1)
  subject2 <- as.factor(subject2)
  roi1 <- as.factor(roi1)
  roi2 <- as.factor(roi2)
  channel1 <- as.factor(channel1)
  channel2 <- as.factor(channel2)
  task <- as.factor(task)
  bin_time_range <- as.factor(bin_time_range)
  bin_period_range <- as.factor(bin_period_range)
  wavelet_library <- as.factor(wavelet_library)
  wavelet_name <- as.factor(wavelet_name)

  is_intra <- tolower(is_intra) == "true"
  is_shuffle <- tolower(is_shuffle) == "true"
})
```

### Prepare a key for pairs of region of interest, for further grouping

```{r}
data$roi_pair <- paste(data$roi1, data$roi2, sep = "-")
```

\newpage

### Inspect the loaded data

```{r}
str(data)
```

\newpage

### Prepare a subset with relevant data

Remove any intra-subject and shuffled dyad from our data set to be analysed, to have only inter-subject dyads

```{r}
sel_not_intra <- data$is_intra == FALSE
sel_not_shuffle <- data$is_shuffle == FALSE
sel_not_na <- sapply(data$coherence, is.na) == FALSE
data_inter <- subset(data, sel_not_intra & sel_not_shuffle & sel_not_na)
```

```{r}
head(data_inter)
```

\newpage

### Show simple inter-subject coherence matrix

The ordering here is lost compared to python code. We use ggplot2 library.

```{r}
ggplot(data = data_inter, aes(x = channel1, y = channel2, fill = coherence)) +
  geom_tile() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  ggtitle("Inter subject coherence")
```

\newpage

### Show inter-subject coherence matrix per Region of Interest

```{r}
ggplot(data = data_inter, aes(x = roi1, y = roi2, fill = coherence)) +
  geom_tile() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  ggtitle("Inter subject coherence per region of interest")
```

\newpage

### Plot the coherence per task, for each dyad

```{r}
ggplot(data = data_inter, aes(x = coherence, y = dyad, color = task)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  ggtitle("Coherence")
```

\newpage

### Compare the coherence by zone pairs with facets

This is only to give inspiration, there are much better ways to visualize this.

```{r}
ggplot(data = data_inter[data_inter$task != "baseline", ], aes(x = coherence, y = roi1)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  scale_y_discrete(labels = function(x) paste("Parent:", x)) +
  facet_wrap(. ~ roi2, labeller = labeller(roi2 = function(x) paste("Child:", x))) +
  ggtitle("Zone pair coherence")
```

\newpage

### Split of coherence per periods range.

```{r}
ggplot(data = data_inter, aes(x = coherence, y = bin_period_range, color = task)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  ggtitle("Coherence per period range")
```

